Introducing statistics with \\texttt{intRo}

### 2014-12-12 student-paper-competition-2014/paper.Rmd
Formatting of the document

---
title: "Introducing statistics with \\texttt{intRo}"
author: "Eric Hare & Andee Kaplan"
date: "December 14, 2014"
output:
  pdf_document:
    includes:
      in_header: preamble.tex
    number_sections: yes
bibliography: refs.bib
---
\thispagestyle{empty}

@@

# Introduction
The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community. However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student. This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work. Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.

@@

Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application. This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}. \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.

@@

## Easy
In order for an application to be useful to students in their first statistics course, it can't hog the spotlight. Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to learn software. With this in mind, in the development of \texttt{intRo} we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find exactly what they need easily. With this easy to use framework, we also made the functionality available in \texttt{intRo} the minimal necessary for an introductory statistics course. Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student.

>>>>

---
title: "Introducing statistics with \\texttt{intRo}"
author: "Eric Hare & Andee Kaplan"
date: "December 14, 2014"
output:
  pdf_document:
    includes:
      in_header: preamble.tex
    number_sections: yes
bibliography: refs.bib
geometry: margin=0.8in
---
\thispagestyle{empty}

@@

# Introduction

The widespread adoption of \texttt{R} as a tool for statistical analysis has undoubtedly been a great development for the scientific community. However, using \texttt{R} in most cases still requires a basic knowledge of programming concepts which may be unfamiliar to the introductory statistics student. This issue may explain why introductory courses often utilize software applications such as \texttt{JMP}, even though instructors would tend to use a tool like \texttt{R} in their own work. Still, signficant compromises must be made when using software like \texttt{JMP}, including dealing with software licenses, unsupported desktop platforms, and non-ideal user interface decisions.

@@

Upon the release of \texttt{RStudio}'s \texttt{Shiny} [@shiny] it became easy for an R-based analysis to be converted to an interactive web application. This in turn, led to the creation of an introductory statistics application which we call \texttt{intRo}, available at http://www.intro-stats.com. \texttt{intRo} offers a number of key advantages over traditional statistics software, which will be discussed at length in this text.

@@

## Easy

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/user_experience.pdf}
\caption{Typical user experience of generating a result in \texttt{intRo}. In this instance, a mosaic plot (and its corresponding code) was created and saved to the final compendium.}
\label{fig:user_experience}
\end{figure}

In order for an application to be useful to students in their first statistics course, it can't hog the spotlight. Statistics as a subject should shine through as a fun and interesting topic to study, rather than being bogged down by difficult to learn software. With this in mind, in the development of \texttt{intRo} we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find exactly what they need easily. With this easy to use framework, we also made the functionality available in \texttt{intRo} the minimal necessary for an introductory statistics course. Additional functionality is nice, but if it clutters the UI beyond the point of usability, it will frustrate an introductory student. Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in \texttt{intRo}, and view the code producing the result.
